---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# RhodopXin

## Description

`RhodopXin` is an R package designed for the exploratory analysis of rhodopsins by looking at their structural, functional, and sequence characteristics. Its main feature would be integrating homology modelling with motif discovery which would help in identifying functionally important rhodopsin-specific motifs across different species, giving an insight as to how they have evolved across organism.

Currently, there are no pre-existing R tools that directly work with rhodopsins. Moreover, this package would leverage pre-existing bioinformatics tools for protein analysis and extend their usage to something that is more specific i.e. exploring the properties of rhodopsins in a comprehensive manner.


## Installation

To install the latest version of the package:

```{r}
install.packages("devtools")
library("devtools")
devtools::install_github("calicac001/RhodopXin", build_vignettes = TRUE)
library("RhodopXin")
```

To run the shiny App: Under Construction

## Overview

To see an overview of the package, run:

```{r}
ls("package:RhodopXin")
data(package = "RhodopXin") 
browseVignettes("RhodopXin")
```

RhodopXin contains # functions.

1. 

An overview of `RhodopXin` is illustrated below.
TBD

## Contributions

The author of the package is Chloe Nichole Calica. The author wrote the functions.

## References

Silva, A. (2022) TestingPackage: An Example R Package For BCB410H. https://github.com/anjalisilva/TestingPackage

Wickham, H., François, R., Henry, L., Müller, K., Vaughan, D. (2023). dplyr: A Grammar of Data Manipulation. https://dplyr.tidyverse.org

## Acknowledgements

This package was developed as part of an assessment for 2024 BCB410H: Applied Bioinformatics course at the University of Toronto, Toronto, CANADA. `RhodopXin` welcomes issues, enhancement requests, and other contributions. To submit an issue, use the [GitHub issues](https://github.com/calicac001/RhodopXin/issues). Many thanks to those who provided feedback to improve this package.
