---
title: "RhodopXin Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{RhodopXinTutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.height = 5,
  fig.width = 7,
  out.width = "100%"
)
```

## Introduction

`Rhodopxin` is an R package designed for the exploratory analysis and visualization of rhodopsin protein structures, focusing on template-based helix alignments and mapping these alignments onto 3D structures of well-characterized template rhodopsins.

To download **RhodopXin**, use the following commands:

```{r, eval=FALSE}
require("devtools")
devtools::install_github("calicac001/RhodopXin", build_vignettes = TRUE)
```

Use the following command to load and attach the package.

```{r setup}
library(RhodopXin)
```

To list all functions available in the package:
```{r, eval=FALSE}
ls("package:RhodopXin")
```

To list all sample datasets available in the package:
```{r, r, eval=FALSE}
data(package = "RhodopXin")
```

## Loading Input

`RhodopXin` works on the amino acid sequences of two following inputs: 
- template rhodopsin: a well-characterized rhodopsin with a resolved structure
- list of query rhodopsins: the subject of sequence alingment ad structure mapping

`RhodopXin` can take in a FASTA file containing either the sequences of the 
template rhodopsin or the query rhodopsins and process it to be used by the package.
Below is an example of loading the two .fasta files provided in `Rhodopxin`.

### Template Rhodopsins
This file contains five template rhodopsins which are all microbial. The first two 
template rhodopsins were selected from the del Val et al. (2014) study. the first
one is a bacteriorhodopsin while the second one is a chimera for channelrhodopsin 
1 and 2. Additionally, the other three template rhodopsins are a type of halorhodopsin, 
proteorhodopsin and xanthorhodopsin. (See ?template_rhodopsins for more details)
```{r}
# Define the path to the fasta file
templatePath <- system.file("extdata", "templates.fasta", package = "RhodopXin")
template <- loadSequence(templatePath)
template
```

### Sample Query Rhodopsins
This file contains three rhodopsin variant selected from the del Val et al. (2014) 
study  which originally looked at 11 channelrhodopsins in a bioinformatic perspective.
```{r}
# Define the path to the fasta file
samplePath <- system.file("extdata", "rhodopsins.fasta", package = "RhodopXin")
sample <- RhodopXin::loadSequence(samplePath)
sample
```
Alternatively, sequences of template rhodopsins can be retrieved just by providing
a valid RCSB PDB id and it will be processed the same way as above. Note that extra
information is printed by the ***read.pdb*** function which is used internally
to fetch the PDB structure from the RCSB database.
```{r}
# Same RCSB id as template 2 above
template2 <- loadFromRCSB("3UG9")
template2
```

### Helix Alignments

This is the main feature of `RhodopXin`. The function ***createHelixAlignments***
performs pairwise alignments between each of the template rhodopsin's helices and 
each of the query sequences, allowing users to see the conservation of residues 
across a given helix.

To see an example, we use the second template rhodopsin and query it against the
sample rhodopsins we loaded above. Note that many messages may come from the external
packages used by this function. It is normal to see 'GET' requests, PDB files in
templorary directories and and alignment messages.
```{r}
alignments <- createHelixAlignments(template = template2, 
                                    sequences = sample, 
                                    rcsb_id = "3UG9")

```
### Output 1: all_pwa
This output is a list of AAStringSet objects containing the combined pairwise alignments
of each template's helix against all the given queries.
```{r}
alignments$all_pwa
```

### Output 2: template_ranges
This output is a dataframe containing information on which positions of the query 
sequence were aligned to the helix. This will still need to be processed in order
to be mapped into the 3D structure.
```{r}
alignments$template_ranges
```
## Visualize Outputs
There are two ways of visualizing the outputs of ***createHelixAlignments***.

### Visualize Helix Alignments
This function combines all of the helices alignments in one plot. For each helix plot,
we have the helix sequence of the template aligned with the query sequences. The 
left hand side of this plot shows which residue number this alignment comes from 
as well as which query. Additionally, a sequence logo and a consensus plot of the 
alignment is present on the top and bottom of the plot.

Note the messages printed are from the external packages used. Also, warning is 
suppressed in this chunk because the plotting function used keeps throwing a 
warning although plots are unaffected. Lastly, the figures may appear squished
but this is due to the rendering of the markdown. Thus, it is recommended to run
this function in the R Shiny app because there, the size is configured for any input.
```{r, warning=FALSE, fig.height=10, fig.width=8}
visualizeHelixAlignments(alignments$all_pwa)
```

### Visualize Helix Mapping
This function maps the conserved residues of a given query into the 3D structure 
of the given template rhodopsin. Here we choose the first query to map into the
structure.
```{r}
visualizeHelixMapping(template_ranges = alignments$template_ranges, 
                      template = template2, 
                      rcsb_id = "3UG9", 
                      query_num = 1)
```


## References
del Val, C., Royuela-Flor, J., Milenkovic, S., & Bondar, A.-N. (2014).
Channelrhodopsins: A bioinformatics perspective. Biochimica et Biophysica Acta
(BBA) - Bioenergetics, 1837(5), 643–655. https://doi.org/10.1016/j.bbabio.2013.11.005

## Session Information

```{r}
sessionInfo()
```

